# Problemset Example
Example of some features of an RTutor interactive Problemset
Author:  Sebastian Kranz

#< ignore
```{r "create_ps"}
# Run all code in this chunk to create and show
# the problem set

library(RTutor)
library(yaml)
# set to TRUE to facilitate error detection
restore.point.options(display.restore.point=FALSE)

# Adapt to your working directory 
setwd("/home/rtutor/ps/Example")

# Specify problem set name
ps.name = "Example"
sol.file = paste0(ps.name,"_sol.Rmd")

# character vector of all packages you load in the problem set
libs = c("leaflet") 

# Compile the problem set
create.ps(sol.file=sol.file, ps.name=ps.name, user.name=NULL,libs=libs, stop.when.finished=FALSE, addons="quiz")

# Show problem set in browser
show.shiny.ps(ps.name, sample.solution=FALSE, launch.browser=TRUE)
stop.without.error()
```
#>

## Exercise 1 -- Summary statistics

We often want to compute some summary statistic of a vector.
For example computing the sum $\sum_{i=1}^n x_i$ of a vector x:

```{r "1 a)",fig.height=3}
#< task
x = 10:20
# Computing the sum of x
sum(x)
#>
```


Now compute the mean of x.
```{r "1 a) 2"}
mean(x)
#< hint
cat("There already exist R functions for many things. To find them use Google, e.g. search for 'R compute mean'.")
#>
```

#< award "mean means mean"
Well, in some occasions one can just guess the name of an R function. The function to compute the mean of a vector, or matrix is called 'mean'. Usually, it is much quicker to goggle than to guess function names, however.
#>

#< info "useful functions for numeric vectors"
Here are examples for useful R functions
```{r "1 "}
max(c(1,5,2)) # returns maximum
min(c(1,5,2)) # returns minimum

sum(c(1,5,2,NA), na.rm=TRUE) # returns sum of all numbers, ignore NA
cumsum(c(1,5,2)) # returns cummulated sum
diff(c(1,5,2)) # returns the vector of differences
```
#>



## Exercise 2 -- Quiz

### A single choice quiz

#< quiz "single"
question: What is 20*20?
sc:
    - 100
    - 200
    - 400*
    - 500
success: Great, your answer is correct!
failure: Try again.
#>

### A multiple choice quiz

#< quiz "multi"
question: Check all prime numbers
mc:
    - 2*
    - 3*
    - 4
    - 5*
    - 6
    - 7*
    - 8
success: Great, all answers are correct!
failure: Not all answers correct. Try again.
#>


### An open question

#< quiz "music"
question: Johann ???? Bach
answer: Sebastian
#>

### An open question with a numeric answer

#< quiz "pi"
question: 2. State pi up to 2 digits
answer: 3.14
roundto: 0.01
#>


### A quiz consisting of several questions (and a single check button)

#< quiz "parts"
parts:
  - question: 1. What is 20*20?
    choices:
        - 100
        - 200
        - 400*
        - 500
    multiple: FALSE
    success: Great, your answer is correct!
    failure: Try again.
  - question: 2. State pi up to 2 digits
    answer: 3.14
    roundto: 0.01
#>


#< award "Quiz master"
You solved the quiz!
#>


## Exercise 3 -- Table outputs

The shiny based inteface of RTutor shows data.frames by default as html tables. This example illustrates how, you can adapt the settings of how data.frames are shown.

Use default settings for printing data.frames, that may have been adapted in the call to `show.ps`.
```{r "mychunk", optional=TRUE}
#< task_notest
T = 40
data.frame(x=runif(T),y=runif(T)*100000)
2*2
#>
```

You can overwrite the default settings for specific chunks, by specifying chunk options.
```{r "mychunk2", signif.digits=5, round.digits=3, table.max.rows=30, html.data.frame=FALSE, optional=TRUE}
#< task_notest
T = 40
data.frame(x=runif(T),y=runif(T)*100000, z=runif(T))
#>
```


## Exercise 4 --Leaflet

Check to see an interactive leaflet centered on the Department of Mathematics and Economics at Ulm University. The place where RTutor has been created.
```{r results='asis', output="htmlwidget", widget="leaflet"}
#< task_notest
SFmap <- leaflet() %>% 
  addTiles() %>% 
  setView(9.9620289,48.4251394, zoom = 11) %>% 
  addMarkers(9.963614702355699,48.42517242865929, popup = 'RTutor was created here at Ulm University')
SFmap
#>
```

Here you can see what else you can do with the great leaflet packages:

http://rstudio.github.io/leaflet/

Here is a list of some other htmlwidgets, which now can also be used with RTutor:

http://www.htmlwidgets.org/
