# Problemset Example


```{r 'check_ps', include=FALSE}

user.name = '' # set to your user name

library(RTutor)
check.problem.set('Example', ps.dir, ps.file, user.name=user.name, reset=FALSE)

# Run the Addin 'Check Problemset' to save and check your solution
```

Example of some features of an RTutor interactive Problemset
Author:  Sebastian Kranz


## Exercise 1 -- Summary statistics

We often want to compute some summary statistic of a vector.
For example computing the sum $\sum_{i=1}^n x_i$ of a vector x:

```{r "1",fig.height=3}
x = 10:20
# Computing the sum of x
sum(x)
```


Now compute the mean of x.
```{r "1__2"}

```


info("useful functions for numeric vectors") # Run this line (Strg-Enter) to show info



## Exercise 2 -- Quiz

### A single choice quiz

#! addon__quiz__single

### A multiple choice quiz

#! addon__quiz__multi


### An open question

#! addon__quiz__music

### An open question with a numeric answer

#! addon__quiz__pi


### A quiz consisting of several questions (and a single check button)

#! addon__quiz__parts




## Exercise 3 -- Table outputs

The shiny based inteface of RTutor shows data.frames by default as html tables. This example illustrates how, you can adapt the settings of how data.frames are shown.

Use default settings for printing data.frames, that may have been adapted in the call to `show.ps`.
```{r "3",optional=TRUE}
T = 40
data.frame(x=runif(T),y=runif(T)*100000)
2*2
```

You can overwrite the default settings for specific chunks, by specifying chunk options.
```{r "3__2",signif.digits=5, round.digits=3, table.max.rows=30, html.data.frame=FALSE, optional=TRUE}
T = 40
data.frame(x=runif(T),y=runif(T)*100000, z=runif(T))
```


## Exercise 4 --Leaflet

Check to see an interactive leaflet centered on the Department of Mathematics and Economics at Ulm University. The place where RTutor has been created.
```{r "4",results='asis', output='htmlwidget', widget='leaflet'}
SFmap <- leaflet() %>% 
  addTiles() %>% 
  setView(9.9620289,48.4251394, zoom = 11) %>% 
  addMarkers(9.963614702355699,48.42517242865929, popup = 'RTutor was created here at Ulm University')
SFmap
```

Here you can see what else you can do with the great leaflet packages:

http://rstudio.github.io/leaflet/

Here is a list of some other htmlwidgets, which now can also be used with RTutor:

http://www.htmlwidgets.org/
